### 2022-05-30 paper.Rmd
more discussion

In this paper, I presented the k-plus criterion for anticlustering. K-plus anticlustering focuses on dividing a data set into subgroups in such a way that these groups are similar to each other. The k-plus criterion extends the classical k-means criterion, which only represents how similar groups are with regard to their means. The k-plus criterion however may be used to also equalize the variance between groups, and any higher order moments such as skewness and kurtosis, as even covariances. Interestingly, k-plus basically does not change the original k-means criterion because it can be reduced to an augmentation of the input data.

A simulation study and examples on real norming data showed that the k-plus criterion is well suited to maximize between-group similarity with regard to multiple criteria. In particular, this work indicates that k-plus anticlustering should be a replacement of the classical k-means approach when striving for between-group similarity. K-means anticlustering disregards any distribution characteristics other than the mean; k-plus anticlustering ensures between-group similarity with regard to multiple criteria while still taking into account differences in means. For large data sets, there is virtually no reason to use k-means anticlustering because the practical differences between k-means and k-plus regarding similarity in means vanish completely. A reasonable exception is given when the input data consists of binary variables because in this case the mean and the standard deviation are directly related.

The results showed that fulfilling several criteria at the same time may be difficult to achieve for small data sets. Still, users can try out different objectives and see which specification fulfils their needs. A reasonable approach it to start greedy, i.e., requesting multiple distribution moments to be equalized, and only if the results are subpar, go for fewer criteria. Unlike in statistical analysis, there is no problem with "just trying out" different approaches when it comes to anticlustering.

K-plus anticlustering was also shown to outperform diversity anticlustering with regard to the specific objectives that are included in the optimization process; diversity is still a reasonable all-arounder that strives for overall betwee-group similarity. Which method should be employed depends on a user's preferences. In doubt, several anticlustering methods can be applied sequentially and the results can be compared directly.

@@

This paper focused on the presentation of new objective function but the did not properly consider the multicriterion nature of the k-plus objective. Instead, a simple (unweighted) sum approach was applied to optimize the k-plus objective, independent of the number of criteria that were incorporated. Other approaches, such as direct algorithms, may yield improved results and better possibilities to select partitionings according to differently weighted criteria. @brusco2019 showed how direct algorithms can be used to optimize multiple critera in anticlustering applications. Their study focused on an approach to find the pareto set of solutions that simultaneously optimized diversity and dispersion, thereby maximizing within-group heterogeneity with regard to both criteria.[^inanticlust]

@@

Focusing on the objetive function only and not on specialized algorithmic optimization approaches followed two reasonings. First, it was of interest to investigate the value of the k-plus objective function per se, independently from the algorithm that is used to optimize it. Hence, to obtain a fair assessment, it was important in the simulation study that each anticlustering objective is optimized using the same procedure, instead of using specialized multicriterion optimization schemes. It was of particular interest to compare the k-plus objective with the popular diversity objective. @papenberg2020 compared k-means anticlustering and diversity anticlustering[^clusterediting] in a simulation study and concluded that the diversity criterion should usually be preferred because k-means does not equalize the spread of the distribution between groups, whereas the diversity maintains an appropriate balance between location and spread. Using the k-plus extension, however, the k-means criterion regains a lot of attractivity: when optimizing for similarity in means and variance, k-plus anticlustering clearly outperformed the diversity criterion in the simulation study. Second, the applications show that the simple sum approach using a general local maximum search yields very satisfying results in practice. It seems that the space of solutions in anticlustering settings includes partitionings that fulfil multiple objectives and these partitionings can be found using a local maximum search.

>>>>

In this paper, I presented the k-plus criterion for anticlustering, which can be used to divided a data set into subgroups in such a way that the groups are similar to each other. The k-plus criterion extends the classical k-means criterion, which only represents how similar groups are with regard to their means. The k-plus criterion also equalizes the variance between groups, and if needed any higher order moments such as skewness and kurtosis, or even the covariances. Thus, k-plus represents a family of anticlustering criteria, which can be combined by users according to their needs. Interestingly, all k-plus variations basically do not change the original k-means criterion, because additional criteria can be incorporated by augmenting the input data.

A simulation study and examples on real norming data showed that the k-plus criterion is well suited to maximize between-group similarity with regard to multiple criteria. In particular, this work indicates that k-plus anticlustering should be a replacement of the classical k-means approach when striving for between-group similarity.[^binarydata]  K-means anticlustering disregards any distribution characteristics other than the mean, while k-plus anticlustering ensures between-group similarity with regard to multiple criteria---and still minimizes differences in means quite well. The results showed that fulfilling several criteria at the same time may be aggregated in small data sets. Still, users can try out different objectives and see which specification fulfils their needs. A reasonable approach it to start greedy by requesting that multiple distribution moments should equalized. Only if the results of the greedy setup are subpar, users can reduce their requirements and opt for fewer criteria. Unlike in statistical analysis, there is no problem with "just trying out" different angles when it comes to anticlustering.

[^binarydata]: A reasonable exception is given when the input data consists of binary variables because in this case the mean and the standard deviation are directly related.

@@

This paper focused on the presentation of new objective function but the did not really consider the multicriterion nature of the k-plus objective. Instead, a simple (unweighted) sum approach---combined with a generic local maximum neighbourhood search---was used to optimize the k-plus objective. Other approaches, such as direct algorithms, may yield improved results and better opportunities to select among partitionings according to multiple differently weighted k-plus criteria. @brusco2019 showed how direct algorithms can be used to optimize multiple critera in anticlustering applications. Their study focused on finding the pareto set of solutions that simultaneously optimize diversity and dispersion -- convincingly arguing that maximizing within-group heterogeneity should incorporate both of these criteria.[^inanticlust] Future research should also investigate how k-plus anticlustering may profit from more sophisticated algorithmic treatment.

@@

Focusing on the objective function only, and not on specialized algorithms, followed two reasonings. First, the applications showed that the local maximum search---maximizing an unweighted k-plus sum---yields satisfying results in practice that probably suffice most users. Second, it was of interest to investigate the value of the k-plus objective function per se, independently from the algorithm that is used to optimize it. Hence, to obtain a fair assessment, it was important in the simulation study that each anticlustering objective is optimized using the same procedure, instead of using specialized multicriterion optimization schemes tailored to each individual criterion. It was of particular interest to compare the k-plus objective with the popular diversity objective. @papenberg2020 compared k-means anticlustering and diversity anticlustering[^clusterediting] in a simulation study and concluded that the diversity criterion should usually be preferred because k-means does not equalize the spread of the distribution between groups, whereas the diversity maintains an appropriate balance between location and spread. Using the k-plus extension, however, the k-means criterion regained attractivity: when opting for similarity in means and variances, k-plus anticlustering clearly outperformed the diversity criterion.
